### 2020-11-05 ijrr.tex
Add abstract.

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[Afour,sageh,times,doublespace]{sagej}

%\usepackage{tabularx}
\usepackage{moreverb,url}
\usepackage{xspace}
%\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{yfonts}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,
  urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\input{new-command}
\newtheorem{definition}{Definition}

\def\volumeyear{2020}

\begin{document}

\runninghead{F.~Lamiraux et al}

\title{Prehensile Manipulation Planning: Modelling, Algorithms and Implementation}

\author{Florent Lamiraux\affilnum{1}, Joseph Mirabel\affilnum{1}}

\affiliation{
  \affilnum{1}LAAS-CNRS, University of Toulouse, France\\
}

\corrauth{Florent Lamiraux, LAAS-CNRS,
  7 avenue du Colonel Roche,
  Toulouse, France
}

\email{florent.lamiraux@laas.fr}

\begin{abstract}
  This paper presents a software platform taylored for prehensile manipulation
  planning named Humanoid Path Planner. The platform implements an original way
  of modelling manipulation planning through a constraint graph that represents
  the numerical constraints that define the manipulation problem. We propose an
  extension of RRT algorithm to manipulation planning that is able to solve a
  large variety of problems. We provide replicable experimental results via a
  docker image that readers may download to run the experimental results by
  themselves.
\end{abstract}

\keywords{robotics, manipulation planning, constrained path planning,
  path planning, software platform}

\maketitle

\input{hpp}

\begin{acks}
This work has been partially supported by Airbus S.A.S. in the framework of the common laboratory Rob4Fam.
\end{acks}

%\bibliographystyle{authordate1}
\bibliographystyle{abbrvnat}

\bibliography{hpp}

\end{document}