### 2021-09-08 review/reply-to-reviewers.tex
Add answer to reviewer comments.

\documentclass{article}
\title{Paper IEEE-TRO 21-0184: reply to reviewers}
\author{Florent Lamiraux and Joseph Mirabel}
\date{}
\begin{document}
\maketitle
\subsection*{Reviewer 7}
\begin{enumerate}
  \item \textbf{Comment.} The beginning of the main content of the
    paper seems hard to follow for readers not familiar with
    manipulation planning.  I think this could be easily improved by
    adding one or two explanatory images earlier on in the
    paper.\\
    \textbf{Answer.} We agree that the first picture comes very lately in the
    initial version of the paper. To improve the paper, we have moved Figure 2
    at the beginning of the paper to provide quickly an example of manipulation
    planning problem.
\end{enumerate}

\subsection*{Reviewer 9}
\begin{enumerate}
\item \textbf{Comment.} In the modeling of SE(2) and SO(2), why using 2
  coordinates for the angular position (sin and cos)? In 2D, using just one
  coordinate (the unique rotation angle) is not enough?\\
  \textbf{Answer.} Using $(\sin\theta,\cos\theta)$ makes the parameterization
  continuous. We added a comment in the paper.

\item \textbf{Comment.} A bit after equation (3), the unnumbered
  equation $q(t)=\cdots$ is the minus sign correct? Should it be the
  operator defined in equation (3)?
  \textbf{Answer.} You are perfectly right. We fixed the error.

\item \textbf{Comment.} I think that some words are supposed to be
  hyperlinks, like ``here''. They do not work.\\
  \textbf{Answer.} You
  are right, there are many hyperlinks in the paper. We checked them
  one by one and found that they all work.

\item \textbf{Comment.}  Algorithm 1: Line 6 checks whether any output
  variable of E is already computed by a previous explicit
  constraint. What if two explicit constraints are redundant? For
  instance, if we have two constraints that (in addition to other
  configuration variables) set the same identical constraint to the
  same config variable.  \\
  \textbf{Answer.} As stated at the beginning of Paragraph IV.D.a, "when an
  explicit constraint is not successfully added following Algorithm 1, it is
  handled as an implicit constraint."

\item \textbf{Comment.} what happens if you detect that a projected
  path is not continuous?\\
  \textbf{Answer.} In this case, only a
  continuous part starting at the beginning of the path is used by the
  M-RRT algorithm. We have added a remark in paragraph and in IV.E.1
  Algorithm 4.

\item \textbf{Comment.} Has the final configuration to be fully
  described? E.g., can I provide just a final desired object pose in
  the cartesian space but not specify the corresponding robot inverse
  kinematic solution (that could be not unique)?\\
  \textbf{Answer.}
  The M-RRT algorithm as of today is not able to define the goal state
  by some constraints, although it would not be difficult to modify
  it for that. However, the python bindings enable users to quickly develop
  new algorithms where the goal can be defined by some constraints. In the
  construction set example described in Section VII.B.2) and in Figure 13, the
  goal is a state of the constraint graph.

\item \textbf{Comment.} Section V.A Grasp: the equation at the
  beginning. Should it be $C = C_{r_1} \times \cdots \times C_{r_{nr}} \times
  SE(3)^{no}$? (I added the dots).\\
  \textbf{Answer.} You are perfectly right. We fixed the error.

\item \textbf{Comment.} In equation (13) (and also in other
  equations), is $R^3 \times SO(3)$ the same as $SE(3)$?\\
  \textbf{Answer.} They both represent rigid-body motion but they have different
  Lie group operators. We have added a paragraph after Equation (13) to clarify
  this point.

\item \textbf{Comment.} The constraint that arises from the stable
  contact pose is piece-wise differentiable. While in the previous
  section the constraints are required to be differentiable. Is there
  any issue?\\
  \textbf{Answer.} Piece-wise differentiable is enough. We have replaced all
  instances of "differentiable" by "piece-wise differentiable".

\item \textbf{Comment.} Section D, after a) Number of states, there is an extra :.\\
  \textbf{Answer.} This has been fixed.

\item \textbf{Comment.}   The used PC has 23 MB of RAM and 3GB of cache. Please
  check these numbers.\\
  \textbf{Answer.} This is 32 GB of RAM and 3 GB of cache. This has been fixed.

\item \textbf{Comment.} I think that Figure 9 is not cited in the text.\\
  \textbf{Answer.} This has been fixed.

\item \textbf{Comment.}  Since the paths are continuous, how is the
  collision checking carried out? Moreover, If I well understand,
  during the waypoint transition the collision checker is not
  called. Is it implicitly assumed here that in such transitions there
  is no collision?\\
  \textbf{Answer.} The software platform proposes discretized collision checking
  as well as continuous collision checking. We added a sentence and a reference
  in the description of the framework. Collisions are checked also along
  waypoint transitions. We added the word "almost" at the end of Section VII-B3). Since the sentence was confusing.
\end{enumerate}

\subsection*{Reviewer 16}
\begin{enumerate}
\item \textbf{Comment.} In the literature survey, it may be good to refer to
  the paper by Lozano-Perez and Kaelbling
  (https://lis.csail.mit.edu/pubs/tlpk-iros14.pdf) because it very much is
  related to representing manipulation planning problems with constraint graphs.\\
  \textbf{Answer.} Although we had already cited 2 papers from this team, we
  have added the requested reference.

\item \textbf{Comment.} Towards the end of Section III, the paper says
  "We denote by $iq_i$ and $iv_i$ the indices of joint $i$ ...". But there
  are multiple indices of joint $i$ (namely $nq_i$ and $nv_i$ many). So
  maybe the sentence should say "the starting (or first?) indices of
  joint i">\\
  \textbf{Answer.} You are right. This has been fixed.

\item \textbf{Comment.} At the end of Section IV-B, the paper refers
  to different linesearch methods implemented. It may make sense to
  briefly describe what these different methods do, or give a
  reference next to each, so that the reader can understand what they
  refer to if interested.\\
  \textbf{Answer.} We have added a reference for one of them and explanations
  for the other ones.

\item \textbf{Comment.} In Section IV-E, the paper says the
  constraints are applied only during "path evaluation". It is not
  perfectly clear what is meant by "path evaluation". What does
  "evaluating" a path mean?\\
  \textbf{Answer.} "Path evaluation" means "computing the configuration at a
  given parameter". We have added this definition in the paper.

\item \textbf{Comment.} In Section V, the paper defines the
  manipulation problem, and as admissible configuration properties, I
  wonder if it should also mention being "collision-free" as a
  requirement. It is kind of obvious, but may be good to be
  complete.\\
  \textbf{Answer.} You are right. We added the absence of collision in the
  definition of admissible configurations.

\item \textbf{Comment.} At the beginning of Section V-A, the combined
  configuration space is defined, but it seems to include only the
  first and the last robot ($C_{r1}$ and $C_{rnr}$). Where are the other
  robots in betwwen 1 and nr?\\ \textbf{Answer.} You are right. This has
  been fixed.

\item \textbf{Comment.} In Section V-A, the freedom of a grasp seems
  to be defined only using binary true/false flags. Using this method,
  is it possible to define limits? For example, for a cylinder, the
  gripper may be able to slide up and down along the height of the
  cylinder. Is it possible to represent using this approach, or would
  one need to discretize along the height of the cylinder? It would be
  good to state this explicitly (whether possible or not). (This
  capability [e.g. limits on translation on a table] is also what is
  needed for the "Stable contact pose" formalization in Section
  V-B. This raises the question: is there really a need to have two
  separate formalizations for hand-object and object-surface
  attachments, or can there be one unified
  formalization?)\\
  \textbf{Answer.} Allowing a free translation along the axis of a cylindrical
  object is possible using flags. However, the user needs to manually add
  inequality constraints to limit this translation. We have added this case in
  the paper with an explanation. Limits of translation are automatically
  defined for contact surfaces by providing plane convex polygons. It seems
  to us difficult to merge these two types of constraints.

\item \textbf{Comment.} One nice trick in Simeon et al. [30] is the
  "virtual closed chain" motions when an object is both on support
  surface and is grasped. Is it possible to have similar solutions
  using the formalization here?\\
  \textbf{Answer.} The constraint graph formalism actually enables us to
  generalize the reduction property in [30]. For instance, the benchmark
  "Romeo placard" could be solved this way. To do so, we need to modify the
  transition constraints when building the constraint graph.

\item \textbf{Comment.} In the experimental results section (VII-B) it
  would be good to state the dimensionality of the problems
  (i.e. robot number of DoF and maybe even objects' number of DoF) as
  that would put the planning times a bit more in
  perspective.\\
  \textbf{Answer.} We have added the number of degrees of freedom to each
  benchmark.
\end{enumerate}

\subsection*{Associate editor}
\begin{enumerate}
\item \textbf{Comment.} In addition, given the importance of the
  software framework, the authors are advised to make sure to provide
  a link to the software repositories on github and to give a proper
  visibility in the manuscript.\\ \textbf{Answer.} There was already a link
  to the installation page, but we made it more visible in the introduction.
\end{enumerate}

\item \textbf{Comment.} Don't use citations as nouns like in "evoked
  in [2]". Rather use "evoked by Mirabel~\emph{et al.}~\cite{...}.
  Don't start a sentence with a citation like "[47] propose". This
  would be fixed by the above change.\\
  \textbf{answer.} We changed citations as requested.

\item \textbf{Comment.} 
Overall, this paper is substandard with respect to writing. The authors
are strongly encouraged to carefully revise the writing and spelling of
the paper. If this should not be corrected in the revised version, this
paper will not be recommended for publication despite the contributions
and this publication recommendation. The authors are encouraged to ask
a native speaker for advise.\\ \textbf{Answer.}
\end{document}